---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "80%",
  dpi = 300
)
```

# rangertoolkit

<!-- badges: start -->
[![R-CMD-check](https://github.com/AFC-AI2C/rangertoolkit/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/AFC-AI2C/rangertoolkit/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

`Rangertoolkit` is intended to be a collection of convenient utilities for working with `ranger` random forest models.
It currently provides the `score_oob` function to generate out-of-bag curves for regression, classification, and survival forests.
It uses the `yardstick` package for evaluation in order to integrate nicely with `tidymodels` and `tidyverse` work flows.

The [OOBCurve](https://github.com/PhilippPro/OOBCurve) package is an alternative resource for generating out-of-bag curves for ranger models; however, it is based around the `mlr` package and does not support survival models.

## Installation

You can install the development version of `rangertoolkit` with devtools.

``` r
devtools::install_github("https://github.com/AFC-AI2C/rangertoolkit/")
```

## Examples

### Regression

This is an example of evaluating the out-of-bag error for a regression model.

```{r example}
library(rangertoolkit)
library(ggplot2)
set.seed(0)

regression_model <- ranger::ranger(
  mpg ~ .,
  data = mtcars,
  num.trees = 200,
  keep.inbag = TRUE
)
regression_score <- score_oob(
  regression_model,
  mtcars[, -1],
  mtcars[, 1]
)

print(head(regression_score))
```

```{r regression_plot}
ggplot(regression_score,
       aes(x = num_trees,
           y = .estimate)) +
  geom_point() +
  geom_line() +
  facet_grid(rows = vars(.metric),
             scales = "free_y") +
  ggtitle("Regression Score") +
  xlab("Number of Trees") +
  ylab("Metric") +
  theme_bw()
```

### Classification

This is an example of evaluating the out-of-bag error for a classification model.

```{r}
classification_model <- ranger::ranger(
  Species ~ .,
  data = iris,
  num.trees = 200,
  keep.inbag = TRUE
)
classification_score <- score_oob(
  classification_model,
  iris[, -5],
  iris[, 5]
)

print(head(classification_score))
```

```{r classification_plot}
ggplot(classification_score,
       aes(x = num_trees,
           y = .estimate)) +
  geom_point() +
  geom_line() +
  facet_grid(rows = vars(.metric),
             scales = "free_y") +
  ggtitle("Classification Score") +
  xlab("Number of Trees") +
  ylab("Metric") +
  theme_bw()
```

### Survival

This is an example of evaluating the out-of-bag error for a survival model. 
This might take some time to run for models with a large number of unique event times.

```{r}
lung_clean <- na.omit(survival::lung)
survival_model <- ranger::ranger(
  survival::Surv(time, status) ~ .,
  data = lung_clean,
  num.trees = 500,
  keep.inbag=TRUE
)
survival_score <- score_oob(
  survival_model,
  lung_clean[, -c(2, 3)],
  survival::Surv(lung_clean$time, lung_clean$status)
)

print(head(survival_score))
```

```{r survival_plot}
ggplot(survival_score,
       aes(x = num_trees,
           y = 1 - .estimate)) +
  geom_point() +
  geom_line() +
  ggtitle("Survival Score") +
  xlab("Number of Trees") +
  ylab("1 - C-Index") +
  theme_bw()
```

